---
title: "Tutorial: TACC Stampede2"
author: "Sangdon Lim [&lt;sangdonlim@utexas.edu&gt;](mailto:sangdonlim@utexas.edu)"
date: "Quantitative Methods<br/>Department of Educational Psychology<br/>University of Texas at Austin"
output:
  revealjs::revealjs_presentation:
    theme: night
    highlight: zenburn
    transition: none
    self_contained: true
    css: slides.css
    reveal_options:
      slideNumber: "c/t"
---

# Supercomputer

* A supercomputer is made up of many computers called "nodes"
* Each node has some number of "workers"
* A desktop computer (1 node) typically has 4 - 8 workers (threads)
* Run this in `R` to find out the # of workers on your machine:

```
parallel::detectCores()
```

* TACC Stampede2 has 4200 nodes * 272 workers

# Parallelization

* Involves distributing a set of tasks to workers
* 1 task in Monte Carlo simulation:
  * processing one replication of a single condition, from data generation to analysis

* 16 conditions * 100 reps = 1600 tasks

* Each worker performs:
  1) Process the allocated task
  2) Returns the result to the task manager
  3) Request another task
  4) Repeat 1-3 until completed

# Converting your code

```{r, eval = F}
# The non-parallel version







n_tasks <- 100
results <- matrix(NA, n_tasks, 2)
for     (task in 1:n_tasks                  )         {
  set.seed(task)
  tmp <- runif(1, 1, 2)
  Sys.sleep(tmp)
  results[task, ] <-
  c(task, tmp)
}
```

# Converting your code

```{r, eval = F}
# The non-parallel version with foreach
library(foreach)






n_tasks <- 100
results <-
foreach (task =  1:n_tasks, .combine = rbind) %do%    {
  set.seed(task)
  tmp <- runif(1, 1, 2)
  Sys.sleep(tmp)

  c(task, tmp)
}
```

# Converting your code

```{r eval = F}
# The parallel version (for your computer)
library(foreach)
library(doParallel)

n_cores <- detectCores() - 2 # Reserve a few workers for other tasks
cl <- makeCluster(n_cores)
registerDoParallel(cl)

n_tasks <- 100
results <-
foreach (task =  1:n_tasks, .combine = rbind) %dopar% {
  set.seed(task)
  tmp <- runif(1, 1, 2)
  Sys.sleep(tmp)

  c(task, tmp)
}
```

# Converting your code

```{r, eval = F}
# The parallel version (for Stampede2)
library(foreach)
library(doParallel)
library(Rmpi)

cl <- getMPIcluster()
registerDoParallel(cl)

n_tasks <- 100
results <-
foreach (task =  1:n_tasks, .combine = rbind) %dopar% {
  set.seed(task)
  tmp <- runif(1, 1, 2)
  Sys.sleep(tmp)

  c(task, tmp)
}
```

# Code prep

Organize your simulation like this:

```{r, eval = F}
conditions <- expand.grid(
  IV1 = c(100, 300, 500),
  IV2 = c(1, 2, 3)
)
task_list <- expand.grid(
  idx_condition = 1:dim(conditions)[1],
  idx_trial     = 1:100
)
n_tasks <- dim(task_list)[1]
tasks <- 1:n_tasks
tasks <- sample(tasks)

foreach (task =  1:n_tasks, .combine = rbind) %dopar% {
  idx_condition   <- task_list$idx_condition[task]
  idx_replication <- task_list$idx_replication[task]
  set.seed(idx_replication)
  # do your thing here
}
```

The key is to have one script run everything!

# Accessing TACC

needs `ssh`.

Windows users need some setup for this.

1. Install WSL. Google this!

- (use either WSL1 or WSL2)

2. Install Ubuntu from Windows store.
- (use the most recent version)

3. Run Ubuntu and create your username and password.
- (this stays on your computer)

4. Also install Windows Terminal from Windows store.

5. Use Windows Terminal to use Ubuntu.

# Accessing TACC

needs `ssh`.

Mac users can use Terminal from Launchpad.

# Accessing TACC

Now we have the console ready.

Run this to get in:
```
ssh <<your tacc username>>@stampede2.tacc.utexas.edu
```

example:
```
ssh sangdonlim@stampede2.tacc.utexas.edu
```

Type in your TACC password and the 2FA code.

- TACC has its own 2FA app
- Instructions available on TACC website
- This is different from Duo used on UT websites

If your TACC account does not belong to a faculty project you won't be able to login yet.

# Locale:

If you get error messages about locale after login, it is okay to ignore it.

But if you want to fix it:

1. Run this to exit from TACC:

```
exit
```

2. Run this on your console:

```
sudo update-locale LC_ALL=en_US.UTF-8 LANG=en_US.UTF-8
```

3. Close and reopen the console.

# Using Stampede2

1. Load R module
2. Install packages you need
3. Upload files
4. Submit a job
5. Have a ☕
6. Download files

# Load R module

Run this:

```
module load Rstats/3.5.1
```

Note: you must use R 3.5.1!

R 4.0.3 is on TACC but it does not run in parallel yet (as of March 2022)

# Install R packages

Run this:
```
R
```

and `install.packages()`

Installed packages stay on your TACC account.

To exit from R, run `q()`

# Install R packages

Because R 3.5.1 is old..

`install.packages()` may not work for some packages.

To install old versions manually:

```
library(devtools)
install_version("mirt", version = "1.31", repos = "http://cran.us.r-project.org")
```

# Install R packages

or raise a support ticket on TACC to let them know we need R 4.0.3!

# Upload files

Open another console. Run this:

```
sftp <<your tacc username>>@stampede2.tacc.utexas.edu
```

example:
```
sftp sangdonlim@stampede2.tacc.utexas.edu
```

# Upload files

Use these Linux commands:

```
pwd          ## show the current working directory on remote (Stampede2)
lpwd         ## show the current working directory on local (your computer)

ls           ## list all files in remote working directory
lls          ## list all files in local working directory

cd mydir     ## open a folder named mydir on remote
lcd mydir    ## open a folder named mydir on local (your computer)

mkdir mydir  ## create a folder named mydir on remote
lmkdir mydir ## create a folder named mydir on local

put *        ## upload all files from remote wd to local wd
get *        ## download all files from remote wd to local wd
```

# Upload files

``~`` is a shortcut to your home directory.

To change your remote (TACC) working directory to your remote home directory:

```
cd ~
```

If you are using Windows, your C drive is located on ``/mnt/c/``.

To change your local working directory to your local C drive:

```
lcd /mnt/c/
```

# Submit a job

needs a SLURM script.

Create a file named `run.sh` with 

```
#!/bin/bash
#SBATCH -J sim               # Job name
#SBATCH -o sim.o             # Name of stdout output log file
#SBATCH -e sim.e             # Name of stderr output log file
#SBATCH -N 4                 # Total number of nodes to request
#SBATCH -n 32                # Total number of workers to request (distributed over nodes)
#SBATCH -p development       # The type of queue to submit to
#SBATCH -t 0:10:00           # Time limit to request (hh:mm:ss)
#SBATCH -A YOUR_PROJECT_ID   # Your project name
#SBATCH --mail-user=YOUR@EMAIL.EDU # TACC will send emails with status updates
#SBATCH --mail-type=all            # Get all status updates

# load R module
module reset
module load Rstats/3.5.1

# call R code from RMPISNOW
ibrun RMPISNOW < main.R
```

This job will run `main.R`.

# Submit a job

Now submit the job:

```
sbatch run.sh
```

The job is now on the waiting line.

# Task allocation

In the SLURM script we requested for:

- 32 workers to be distributed over 4 nodes

This does not mean we will get 8 workers in each node!

- worker assignments are done dynamically
- users do not have direct control

# Time limit

The job will run for 10 minutes as specified in the SLURM script.

If your R code does not complete within the time limit:

- TACC will force-stop the code.
- Prepare your code for this!

# Have a ☕

To see the status of the job:

```
showq -u
```

To cancel the job:

```
scancel <JOB_ID>
```

example:
```
scancel 123456
```

# If job fails:

Examine log files:

```
cat sim.o
cat sim.e
```

Filenames are specified in the SLURM script.

# TACC storage types

TACC has multiple storage areas

* ``$HOME`` : 10GB, auto backup, permanent
* ``$WORK`` : 1TB, no backup, permanent
* ``$SCRATCH`` : Unlimited (~30PB), no backup, deleted after 10 days

Linux commands to move to each folder:

```
cd $HOME
cd $WORK
cd $SCRATCH
```

R functions to retrieve file paths:

```
Sys.getenv("HOME")
Sys.getenv("WORK")
Sys.getenv("SCRATCH")
```

# The big red button

When you are done with testing on ``development`` queue

* increase # of nodes and workers in SLURM script
* also increase the time limit
* and submit to the ``normal`` queue
* have a 😴

# TACC queue types

https://portal.tacc.utexas.edu/user-guides/stampede2#table5

Each queue has limits on what each job can request.

* ``development``: 16 nodes, 2 hours
* ``normal``: 256 nodes, 48 hours
* ``large``: 2048 nodes, 48 hours
* ``long``: 32 nodes, 120 hours

Your project has SU credits assigned to it.

Use one node for one hour = costs 0.8 SU

Other types of processors have their own queue type. See the link above.

# More on time limit

If you request for 48 h but it completes in 1 h:

- You will be only charged for 1 hour SU.

\

Reasons to not to specify 99999 nodes * 999 hours:

- You will have to wait longer on the list.

- TACC will "hold" the full amount of SU on your project until it knows how much it needs to charge.

# General guide on coding for TACC

1. Save your results to a file after each task.
- TACC will force-stop your code after your time limit.
- You must specify the time limit in SLURM when submitting your job.

\

2. Make your code be able to skip a task when result files are already there.

# General guide on coding for TACC

3. Spend less time on optimizing your code.
- Time works differently on TACC.
- Your fingers are slower than TACC.
- The time you spend on optimizing can be spent on just running your code on TACC.

\

4. Spend more time on making your code error-free.
- TACC has a waiting list.
- If you wait for 5 days on the list and the code errors in 5 seconds then you won't like it.

# General guide on coding for TACC

5. If you are using ``tidyverse``..
- Consider converting your code to not use it.
- ``tidyverse`` is a big package, and it may make your code take a lot more time to run on TACC.

\

# Thanks!

- slides made with ``revealjs`` in R markdown
- source available on https://github.com/SangdonLim/tacc_r_intro
- last updated March 2022
